---
title: "Practical Machine Learning - Project Writeup"
author: "Jerold Paulson"
date: "Sunday, October 26, 2014"
output: html_document
---
First we will define some functions that will be used to clean up the data
We will remove all columns with NAs present.
```{r cache=TRUE}
filterPml       <- function(x) { read.csv(x, na.strings = c("", "NA", "#DIV/0!") ) }
removeNAColumns     <- function(x) { x[ , colSums( is.na(x) ) < nrow(x) ] }
onlyCompleteRows       <- function(x) {x[,sapply(x, function(y) !any(is.na(y)))] }
```
Now we will read in the data from the csv files already downloaded from the web
```{r, cache=TRUE}
trainData      <- filterPml("pml-training.csv")
testData       <- filterPml("pml-testing.csv")
```
Next, we will remove columns 1,2,5 and 6 : 1,2 and 6 are not predictive variables, 5 is redundant
``` {r,cache=TRUE}
training       <- trainData[,-c(1,2,5,6)]
testing        <- testData[,-c(1,2,5,6)]
```
Now, let's partition the training set data for evaluative purposes into training and testing 70/30
```{r,cache=TRUE}
library(caret)
trainIndex  <- createDataPartition(training$classe, p=.70, list=FALSE)
trainingSubset <- training[ trainIndex,]
testingSubset  <- training[-trainIndex,]
```
Now filter out the NA Columns / incomplete rows
```{r,cache=TRUE}
trainingSubset <- onlyCompleteRows(removeNAColumns(trainingSubset))
testingSubset  <- onlyCompleteRows(removeNAColumns(testingSubset))
```
Next we do the actual training and look at the accuracy on the test subset of the practice data
```{r, cache=TRUE}
library(randomForest)
library(e1071)
boost <- train(classe~.,data=trainingSubset,method="gbm",verbose=FALSE,trControl=trainControl(method="cv",repeats=10))
```
Let's now look at the results of the analysis and the quality of the predictions
``` {r}
print(summary(boost))
print(confusionMatrix(predict(boost,newdata=testingSubset[,-56]),testingSubset$classe))
```
Lastly, we find the actual predictions for the 20 test cases used for the assignment
``` {r}
answer <- predict(boost, newdata=testData)
print(answer)
```